---
title: "Assignment6"
author: "Junyi_Liu"
date: "2022-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this Assignment, we will compare the three methods to generate the random number in two ways: sampling time and quality of the sample.
```{r set seed}
set.seed(42)
```
Firstly, we will use 'rnorm' method and repeat the process 50 times.
```{r rnorm}
delta_t_list_rnorm = c()
KS_stats_list_rnorm = c()
for (i in seq(1,50,by = 1)){
t0 = Sys.time()
X <- rnorm(10^5)
t1 = Sys.time()
delta_t = t1-t0
delta_t_list_rnorm<-append(delta_t_list_rnorm,delta_t)

F_hat <- ecdf(X)
F_hat_eval <-F_hat(sort(X))
F_eval <- pnorm(sort(X))
KS_stats <-max(abs(F_hat_eval- F_eval))
KS_stats_list_rnorm<-append(KS_stats_list_rnorm,KS_stats)
}
```
Secondly, we will use 'Box-Muler transform' method and repeat the process 50 times as well.(Here we conbine the Z1 and Z2.)
```{r BOX}
delta_t_list_box = c()
KS_stats_list_box = c()
for (i in seq(1,50,by = 1)){
t0 = Sys.time()
U1 = runif(10^5/2)
U2 = runif(10^5/2)
Z1 = sqrt(-2*log(U1))*cos(2*pi*U2)
Z2 = sqrt(-2*log(U1))*sin(2*pi*U2)
t1 = Sys.time()
delta_t = t1-t0
delta_t_list_box<-append(delta_t_list_box,delta_t)

Z = c(Z1,Z2)
F_hat <- ecdf(Z)
F_hat_eval <-F_hat(sort(Z))
F_eval <- pnorm(sort(Z))
KS_stats <-max(abs(F_hat_eval- F_eval))
KS_stats_list_box<-append(KS_stats_list_box,KS_stats)
}
```

Now, for the last method--'rejection sampling' and time consuming here will be much more.
```{r rejection sampling}
delta_t_list_reject = c()
KS_stats_list_reject = c()
for (i in seq(1,50,by = 1)){

M <- optimize(f=function(x){dnorm(x)/dexp(abs(x))*2}, interval=c(0,5),maximum=T)$objective

# returns a single accepted sample from standard doubly exp. proposal and N(0,1) target
acc_rej_dexp_norm <- function(M,a,b){
  U <- runif(1)*M
  sings <- c(-1,1)
  Y <- rexp(1)*sample(sings,1) # doubly exponential
  while (U>dnorm(Y)/dexp(abs(Y))/2/M){
    U <- runif(1)*M
    Y <- rexp(1)*sample(sings,1)
  }
  return(Y)
}

N <- 10^5
X_reject <- rep(0,N)
t0 = Sys.time()
for(n in 1:N){
  X_reject[n] <- acc_rej_dexp_norm(M,a,b)
}
t1 = Sys.time()
delta_t = t1-t0
delta_t_list_reject<-append(delta_t_list_reject,delta_t)

F_hat <- ecdf(X_reject)
F_hat_eval <-F_hat(sort(X_reject))
F_eval <- pnorm(sort(X_reject))
KS_stats <-max(abs(F_hat_eval- F_eval))
KS_stats_list_reject<-append(KS_stats_list_reject,KS_stats)
}
```

```{r}
KS_rnorm = mean(KS_stats_list_rnorm)
KS_box = mean(KS_stats_list_box)
KS_reject = mean(KS_stats_list_reject)

delta_t_rnorm = mean(delta_t_list_rnorm)
delta_t_box = mean(delta_t_list_box)
delta_t_reject = mean(delta_t_list_reject)

```
The sampling time used for 'rnorm' is:

```{r}
print(delta_t_rnorm)
```
The sampling time used for 'Box-Muler transform' is:
```{r}
print(delta_t_box)
```
The sampling time used for 'rejection sampling' is:
```{r}
print(delta_t_reject)
```
To draw a conclusion, for the same sample size,the time used for 'rnorm' is the least,time for 'Box-Muler transform' is the second least and the time for 'rejection sampling' is the most---around 10 seconds!

Now we talk about the quality of the number, identifying by the KS statistic.
KS statistic for the 'rnorm' is:
```{r}
print(KS_rnorm)
```

KS statistic for the 'Box-Muler transform' is:
```{r}
print(KS_box)
```

KS statistic for the 'rejection sampling' is:
```{r}
print(KS_reject)
```

To draw a conclusion, for the same sample size,the quality for the three methods is nearly the same. The smaller the KS statistic is ,the better quality it will have.