---
title: "Assignment6"
author: "Junyi_Liu"
date: "2022-11-18"
output: html_document
---

In this Assignment, we will compare the three methods to generate the random number in two ways: sampling time and quality of the sample.
```{r set seed}
set.seed(42)
```
Firstly, we will use 'rnorm' method and repeat the process 50 times.
```{r rnorm}
delta_t_list_rnorm = c()
KS_stats_list_rnorm = c()
for (i in seq(1,50,by = 1)){
t0 = Sys.time()
X <- rnorm(10^5)
t1 = Sys.time()
delta_t = t1-t0
delta_t_list_rnorm<-append(delta_t_list_rnorm,delta_t)

F_hat <- ecdf(X)
F_hat_eval <-F_hat(sort(X))
F_eval <- pnorm(sort(X))
KS_stats <-max(abs(F_hat_eval- F_eval))
KS_stats_list_rnorm<-append(KS_stats_list_rnorm,KS_stats)
}
```
Secondly, we will use 'Box-Muler transform' method and repeat the process 50 times as well.(Here we conbine the Z1 and Z2.)
```{r BOX}
delta_t_list_box = c()
KS_stats_list_box = c()
for (i in seq(1,50,by = 1)){
t0 = Sys.time()
U1 = runif(10^5/2)
U2 = runif(10^5/2)
Z1 = sqrt(-2*log(U1))*cos(2*pi*U2)
Z2 = sqrt(-2*log(U1))*sin(2*pi*U2)
t1 = Sys.time()
delta_t = t1-t0
delta_t_list_box<-append(delta_t_list_box,delta_t)

Z = c(Z1,Z2)
F_hat <- ecdf(Z)
F_hat_eval <-F_hat(sort(Z))
F_eval <- pnorm(sort(Z))
KS_stats <-max(abs(F_hat_eval- F_eval))
KS_stats_list_box<-append(KS_stats_list_box,KS_stats)
}
```

Now, for the last method--'rejection sampling' and time consuming here will be much more.
```{r rejection sampling}
delta_t_list_reject = c()
KS_stats_list_reject = c()
for (i in seq(1,50,by = 1)){

M <- optimize(f=function(x){dnorm(x)/dexp(abs(x))*2}, interval=c(0,5),maximum=T)$objective

# returns a single accepted sample from standard doubly exp. proposal and N(0,1) target
acc_rej_dexp_norm <- function(M,a,b){
  U <- runif(1)*M
  sings <- c(-1,1)
  Y <- rexp(1)*sample(sings,1) # doubly exponential
  while (U>dnorm(Y)/dexp(abs(Y))/2/M){
    U <- runif(1)*M
    Y <- rexp(1)*sample(sings,1)
  }
  return(Y)
}

N <- 10^4
X_reject <- rep(0,N)
t0 = Sys.time()
for(n in 1:N){
  X_reject[n] <- acc_rej_dexp_norm(M,a,b)
}
t1 = Sys.time()
delta_t = t1-t0
delta_t_list_reject<-append(delta_t_list_reject,delta_t)

F_hat <- ecdf(X_reject)
F_hat_eval <-F_hat(sort(X_reject))
F_eval <- pnorm(sort(X_reject))
KS_stats <-max(abs(F_hat_eval- F_eval))
KS_stats_list_reject<-append(KS_stats_list_reject,KS_stats)
}
```
Here we print the result of KS.
```{r print result for KS}
mean(KS_stats_list_rnorm)
sd(KS_stats_list_rnorm)

mean(KS_stats_list_box)
sd(KS_stats_list_box)

mean(KS_stats_list_reject)
sd(KS_stats_list_reject)
```
To draw a conclusion, for the same sample size,the quality for the 'rnorm' and 'Box-Muler transform' are nearly the same, while the KS of 'rejection sampling' is larger. The smaller the KS statistic is ,the better quality it will have.

Now, we need to print the result of time.
```{r print result for t}
mean(delta_t_list_rnorm)
sd(delta_t_list_rnorm)

mean(delta_t_list_box)
sd(delta_t_list_box)

mean(delta_t_list_reject)
sd(delta_t_list_reject)
```
To draw a conclusion, for the same sample size,the time used for 'rnorm' is the least,time for 'Box-Muler transform' is the second least and the time for 'rejection sampling' is the most---around 2 seconds!